\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-18 arch.tex
updates to arch with few comments

\item \texttt{synchronize} takes the last {\em external} observations
  as an argument and returns the {\em internal} observations for a
  reactor or a failure report.

\item \texttt{step} that accepts new goals requested (if any) on the
  {\em internal} timeline of a reactor and will execute one step of
  deliberation. In the context of \eu for example one deliberation
  step corresponds to a single flaw resolution (Section
  \ref{sec:europa:arch}) while ensuring that the resulting plan is not
  proven inconsistent. This implies that it also includes the
  backtracking of the search in such a case.  The returned value by
  the \texttt{step} function is a boolean which indicates whether the
  reactor needs an extra step to deliberate or if this last step
  resulted in a complete plan within the reactor's look-ahead. On the
  later, the part of the plan that concern this reactor {\em external}
  timelines is provided in order to be dispatched as goals to
  whichever reactor that is reponsible for these. % \comment{By
    % extending I may have saud things I developed below ... will have
    % to check}
\end{itemize}

Both these deliberative functions have different scope and temporal
constraints. The \texttt{synchronize} call is considered as atomic and
will focus only on state inference for the current tick by identifying
the current value of {\em internal} state variables from the latest
updates on the {\em external} state variables. This process propagates
up in the reactor dependency graph at every tick to ensure that all
reactors are aware of the current state of the world. \texttt{step}
relates to a single decision step of the planning process which is
focused on the future evolution of state variables; it can take
multiple steps spanning multiple ticks for a reactor to produce a plan
that is both complete and valid. How the reactor identifies that its
plan is complete and valid is its responsibility and depend on the way
such reactor resolve its plan during each {\em steps}. The general
idea is that the reactor was able to find a valid strategy tin order
to satisfy as much of its {\em internal} goals as possible. In the
context of \eu, which we are going to develop, this means that the
solver do not identify any more flaw to be resolved in the plan.

@@

The manner in which the agent interleaves planning \texttt{step}s and
synchronization for a single reactor is illustrated in
Fig.~\ref{fig:tick-exec}. Synchronization occurs at every tick while
interrupting planning; this allows a reactor to identify its current
state that can propagate through the reactor hierarchy in order to
ensure a consistent view of the state of the world between all the
reactors.
% This view of a reactor has two parallel tasks allowing for long
% planning deliberation while ensuring that the state evolution
% propagates through the agent as time advance.
These intertwined tasks together provide two natural information flows
(bottom-up for state estimation and top-down for plan projection) that
we expect to see in a control loop. As they are both inference-based
processes it is natural to implement a reactor based on generic
automated planning. For example in Fig. \ref{fig:agent}, apart from
\textsf{Shore comm.} and \textsf{Vehicle} acting as interfaces to the
system, every other reactor (in blue) are different instances of a \eu
reactor differentiated only by their model and \eu solver
configurations.

@@

Within \rx, an instance of an \eu reactor comes with the sole purpose
to deliberate. A typical design of such a reactor takes its cues from
\texttt{IDEA} \cite{mus02, mus06} where both planning and execution
are based on a single unified model within a rich representation that
provides support for temporal inference. In \rx, the execution cycle
consists of two deliberation processes which are implemented as two
\eu solvers modifying the \emph{same} plan database. %  and for which the
% execution is managed by the \rx framework as time advance during the
% execution of the system:

@@

\item The \emph{synchronization solver} is a specialized \eu solver
  that integrates new information in the plan, the evolution of
  \emph{external} state variables and ensures that the reactor
  propagates these to identify the reactors current state as well as
  to inform other reactors of any state change on its \emph{internal}
  timelines. This solver is summoned at the beginning of every single
  tick.

\item The \emph{deliberation solver} manages the deliberation process
  of the reactor either to produce a new plan or alter its current
  plan as new goals are provided or when the synchronization solver
  identifies a conflict between current state and expectations from a
  previously generated plan. This process can span multiple ticks and
  can therefore be interrupted by the synchronization solver.

@@

While the two processes are separate, they share the same plan
internal to a reactor; consequently at every synchronization cycle the
planning process is informed of the new world state and its impact on
the plan. Conversely, when synchronization occurs after deliberation
steps, its starting point is the last partial plan resulting from
deliberation. This gives an initial frame to the problem that helps
-- at least initially -- to target the synchronization problem by
ensuring that the new {\em external} state updates are compatible with
this plan. If so, the synchronization propagates these events in the
plan and can uses it as a help on identifying its {\em internal}
timelines based on this plan. If they are not compatibles, then
{\em synchronization} identifies a plan inconsistency  reflecting that the
current plan is not valid anymore. This results on the retraction of
decisions made during {\em deliberation} steps in order to allow 
{\em synchronization} to identify its {\em internal}  state. Such
relaxation can be viewed as a backtracking on the plan search
triggered exogenously by the world evolution. It is the way our
reactor will identify that its current plan failed to execute and can
trigger further deliberations steps in order to find an alternate 
strategy.

Finally, when the deliberation process did identify a complete
solution plan for its current {\em internal} goals. The part of its
plan that applies to its {\em external} timelines is sent as a 
request to \rx that will handle their dispatching to the corresponding 
reactors. If things go smoothly the {\em synchronization} process will
eventually receive {\em external } observations that confirm the
execution of this plan and will allow the plan execution to progress
inside the reactor (by having the part of the plan observed restricted
to the start time of these observations).

Synchronization offers an important challenge for embedding planning
in a situated agent, typically %  as always been to go around the asumptions made by
% most classical planners. Specifically one
an assumption to restrict the problem to ``offline planning'' defined
in \cite{ghallab04} as follows:

\begin{quotation}
  The planner is not concerned with any change that may occur in
  $\Sigma$ \footnote{The world modeled by the plan domain.} while it is
  planning; it plans for the given initial and goal states regardless
  of the current dynamics, if any. 
\end{quotation}

This assumption, while helpful to reduce the scope of the planning
problem, becomes problematic for planning within a highly dynamic
environment. In a typical robot this implies that the only point where
the planner can be informed about the current world state is prior to
the commencement of a planning cycle by giving to the planner the
initial state and current objectives. During the planning, it is
assumed that the agent can maintin the world as perceived by the
planner as stable. \cite{lemai04, lemai-chenevier2004} attempts to
reduce the impact of planning by allowing local plan repair in the
existing plan. In the eventuality that such repair is not feasible
they require that their robot comes to a full stop until (re-)planning
is complete.% \comment{I may need more examples/refs here}. 
As Fig. \ref{fig:tick-exec} shows, this assumption is problematic for our
architecture. Synchronization that tracks state evolution can occur
several times during the reactor deliberation process. By making
synchronization and deliberation share the same plan we provide a
solution that relaxes this static world assumption. According to 
\cite{Dias:2003ua, mus06}, IDEA was sharing the same core idea 
where all the solvers of an agent shared the same plan as a basis of
communication and specifically their {\em Reactive planner} --roughly
similar to our {\em synchronization} solver -- was often the only one 
targeting the execution frontier. Although, none of the papers we
found were detailing how this was handled in the
architecture. Moreover, in \cite{Dias:2003ua} we can read:
\begin{quotation}
  [\textellipsis] as long as the planning horizon over which the
  deliberative planner is working never intersects the current
  execution time, deliberative planning can operate in parallel with
  reactive execution and does not affect the reactivity of the agent. 
\end{quotation}
\comment{fpy: This is my interpretation of this sentence at least .... feel
  free to comment in case you disagree}
Which suggest a non overlap between {\em reactive planner} and {\em
  deliberative planners} horizons. In \rx, and more specifically in
our \eu reactor the planning horizon of {\em synchronization} -- 
limited to the current tick -- is a subset of the planning horizon 
of the deliberation steps -- starting at the current tick and with an
extend based on the latency $\lambda$ and look-ahead $\pi$ of 
the reactor. The way we avoid to create conflict between the two is by
a careful control of the scheduling of these two process by the \rx 
agent.

In the following sections we show how synchronization and deliberation
are implemented independently to show how their interaction through
the plan helps relax this assumption by making planning fully
integrated in the reactor execution cycle while potentially allowing
for more informed planning.

>>>>

\item \texttt{synchronize} takes the last set of {\em external}
  observations as an argument and returns the {\em internal}
  observations for a reactor or a failure report.

\item \texttt{step} that accepts new goals requested (if any) on the
  {\em internal} timeline of a reactor and will execute one step of
  deliberation. In the context of \eu for example, one deliberation
  step corresponds to a single flaw resolution (Section
  \ref{sec:europa:arch}) while ensuring that the resulting plan is not
  inconsistent. Backtracking is included in the step should it be such
  a case.  The returned value by the \texttt{step} function is a
  boolean which indicates whether the reactor needs an extra step to
  deliberate or if this last step resulted in a complete plan within
  the reactor's look-ahead. The part of the plan that covers the
  reactor's {\em external} timelines is provided to be dispatched as a
  goal.
% to whichever reactor that is responsible for these. % \comment{By
    % extending I may have saud things I developed below ... will have
    % to check}

Both these deliberative \kcomment{``deliberative??''; something else
  or remove altogether perhaps?} functions have different scope and
temporal constraints. The \texttt{synchronize} call is considered as
atomic and will focus only on state inference for the current tick by
identifying the current value of {\em internal} state variables from
the latest updates on {\em external} state variables. This process
propagates up in the reactor dependency graph at every tick to ensure
that all reactors are aware of the current state of the
world. \texttt{step} relates to a single decision step of the planning
process focused on the future evolution of state variables; it can
take multiple steps spanning multiple ticks for a reactor to produce a
plan that is both complete and valid. How the reactor identifies that
its plan is complete and valid is its responsibility and depends on
the way each reactor resolves its plan at each {\em step}. The general
idea is that the reactor is able to find a valid strategy in order to
satisfy as many of its {\em internal} goals as possible. For \eu, this
means that the solver is unable to identify any further flaws to be
resolved in the plan.

@@

The manner in which the agent interleaves planning \texttt{step}s and
synchronization for a single reactor is illustrated in
Fig.~\ref{fig:tick-exec}. Synchronization occurs at every tick while
interrupting planning; this allows a reactor to identify its current
state that can propagate through the reactor hierarchy in order to
ensure a consistent view of the state of the world between all the
reactors.
% This view of a reactor has two parallel tasks allowing for long
% planning deliberation while ensuring that the state evolution
% propagates through the agent as time advance.
These intertwined tasks together provide two natural information flows
(bottom-up for state estimation and top-down for plan projection) that
we expect to see in a control loop. As they are both inference-based
processes it is natural to implement a reactor based on generic
automated planning. For example in Fig. \ref{fig:agent}, apart from
\textsf{Shore comm.} and \textsf{Vehicle} acting as interfaces, every
other reactor (in blue) is a different instance of an \eu reactor
differentiated only by their model and \eu solver configurations.

@@

Within \rx, an instance of an \eu reactor comes with the sole purpose
to deliberate. A typical design of such a reactor takes its cues from
\texttt{IDEA} \cite{mus02, mus06} where both planning and execution
are based on a single unified model within a rich representation that
provides support for temporal inference. In \rx, the execution cycle
consists of two deliberation processes which are implemented as two
separate \eu solvers modifying the \emph{same} plan database.

@@

\item The \emph{synchronization solver} is a specialized \eu solver
  that integrates new information in the plan. This signifies the
  evolution of \emph{external} state variables and ensures that the
  reactor propagates these to identify current reactor state as well
  as to inform other reactors of any state change on its
  \emph{internal} timelines. This solver is summoned at the beginning
  of every single tick.

\item The \emph{deliberation solver} manages the deliberation process
  of the reactor. It does so either to produce a new plan, to alter
  its current plan as new goals are provided or when the
  synchronization solver identifies a conflict between current state
  and expectations from a previously generated plan. This process can
  span multiple ticks and can therefore be interrupted by the
  synchronization solver.

@@

While the two processes are separate, they share the same plan
internal to a reactor; consequently at every synchronization cycle the
planning process is informed of the new world state and its impact on
the plan. Conversely, when synchronization occurs after deliberation
stops, its starting point is the last partial plan resulting from
deliberation. This gives an initial frame to the problem that helps to
target the synchronization problem by ensuring that the new {\em
  external} state updates are compatible with this plan. If so,
synchronization propagates these events in the plan and can uses it as
an aid in identifying its {\em internal} timelines \kcomment{``use it
  as an aid to identify'' what though?} based on this plan.

If they are not compatible, then {\em synchronization} identifies a
plan inconsistency. This results in the retraction of decisions made
during the \kcomment{previous?} {\em deliberation} step. % in order to
% allow {\em synchronization} to identify its {\em internal} state.
Such a relaxation can be viewed as a backtracking % on the plan search
% triggered exogenously by the world evolution. It is the way our
allowing the reactor to identify that its current plan has failed to
execute triggering further deliberations steps to find an alternate
strategy.

Finally, when the deliberation process has identified a complete
solution for its current {\em internal} goals, a part of its plan that
applies to its {\em external} timelines is sent as a request to \rx to
handle their dispatch to corresponding reactors.  Synchronization
eventually receives {\em external } observations that confirms plan
execution, in turn recording execution progress by having the part of
the plan observed restricted to observation start times.

Synchronization offers an important challenge for embedding planning
in a situated agent, typically an assumption to restrict the problem
to ``offline planning'' defined in \cite{ghallab04}:

{\scriptsize
  \begin{quote}
    The planner is not concerned with any change that may occur in
    $\Sigma$ \footnote{The world modeled by the plan domain.} while it
    is planning; it plans for the given initial and goal states
    regardless of the current dynamics, if any.
\end{quote}}

This assumption, while helpful to reduce the scope and complexity of
the planning problem, becomes problematic for planning within a highly
dynamic environment. In a typical robot this implies that the only
point where the planner can be informed about the current world state
is prior to the commencement of a planning cycle, by providing the
planner an initial state and current objectives. During the planning,
it is assumed that the agent can maintain the world as perceived by
the planner. \cite{lemai04, lemai-chenevier2004} attempts to reduce
the impact of planning by allowing local plan repair in the existing
plan. In the eventuality that such repair is not feasible they require
that their robot comes to a stop until (re-)planning is complete.

This assumption is problematic for situated agents in general and as
Fig. \ref{fig:tick-exec} shows, for \rx. Synchronization that tracks
state evolution can occur several times during reactor
deliberation. By making synchronization and deliberation share the
same plan we provide a solution that relaxes this static world
assumption. \texttt{IDEA} uses the same principle with all solvers in
an agent sharing the same plan for communicating environmental change
\cite{Dias:2003ua, mus06}. Specifically their {\em Reactive planner}
-- similar to our {\em synchronization} solver -- was often the solver
targeting the change at the execution frontier. 
% Although, none of the papers we found were detailing how this was
% handled in the architecture.
\cite{Dias:2003ua} states:

{\scriptsize
  \begin{quote}
  [\textellipsis] as long as the planning horizon over which the
  deliberative planner is working never intersects the current
  execution time, deliberative planning can operate in parallel with
  reactive execution and does not affect the reactivity of the agent. 
\end{quote}}

This suggests a deliberate non overlap between {\em reactive planner}
and {\em deliberative planners} horizons. In \rx, and more
specifically in our \eu reactor the planning horizon of {\em
  synchronization} -- limited to the current tick -- is a subset of
the planning horizon of the deliberation steps -- starting at the
current tick and with an extent based on latency $\lambda$ and
look-ahead $\pi$ of the reactor. In this manner we avoid creating a
conflict between the deliberation and synchronization by controlling
the scheduling of these two process by the \rx agent.

In the following sections we show how synchronization and deliberation
are implemented independently to show how their interaction through
the plan helps relax this assumption. It does so by integrating
planning in the reactor execution cycle even while allowing for more
informed deliberation.
